Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-07-29 grove-paper.tex
Updates from Overleaf

\subsubsection{Wrapping Code}

\begin{figure}
\centering
TODO: naive add that creates a conflict
\begin{tikzpicture}
\path (0,0)       node {\includegraphics[scale=0.5]{graphviz/wrapping/1.dot}};
\path (0cm,-4cm)  node {\includegraphics[scale=0.5]{graphviz/wrapping/2.dot}};
\path (0cm,-8cm)  node {\includegraphics[scale=0.5]{graphviz/wrapping/3.dot}};
\path (0cm,-12cm) node {\includegraphics[scale=0.5]{graphviz/wrapping/4.dot}};
\path (0cm,-16cm) node {\includegraphics[scale=0.5]{graphviz/wrapping/5.dot}};
\end{tikzpicture}
\caption{Wrapping}
\label{fig:Wrapping}
\end{figure}

Before \marginpar{DEL?} further exploring how to merge changes from multiple users,
we want to point out a common idiom that we call \emph{wrapping}.
Consider the first graph in Figure~\ref{fig:Wrapping}, which
represents the code \texttt{$\lambda$ x . x}, and suppose
the user's cursor is in the body field of the lambda (i.e., around the variable
reference to \texttt{x} in the body of the lambda).

If the user requests that a plus be inserted,
the editor may notice that there is something already in that hole.
A naive editor could certainly add the plus by creating a new
edge from the body index of the lambda to a freshly created \texttt{plus} vertex.
This would result in a conflict between the \texttt{plus} and the existing \texttt{var}
as shown in Figure~REF:TODO.
A better user experience
would be interpret the request for a plus to mean to
take whatever (if anything) is in the cursor
and place it as a child of the plus.
This \emph{wraps} the existing \texttt{var} in the \texttt{plus}
and results in the code \texttt{$\lambda$ x . x + \hole}.
(This is similar to the behavior in other structure editors such as in CITE:TODO.)
Translating this into edge deletes and adds results in the sequence of graphs in Figure~\ref{fig:Wrapping}.
Note that this is a layer above

Traditional: no relation (and indent might change)

\subsubsection{Conflicting wrap operations}

\begin{figure}
\centering
\begin{tikzpicture}
% TODO: texttt
% TODO: position the different graphs
\path (0,0)       node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/1.dot}};
\path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/2A.dot}};
\path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/2B.dot}};
\path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/3.dot}};
% \path (0cm,0cm) graph[tree layout] {
%   lam[as={$\lambda_{\texttt{x}}$}] -> {
%     x
%   }
% };
% \path (-3cm,-2cm) graph[tree layout] {
%   lam[as={$\lambda_{\texttt{x}}$}] -> {
%     + -> {
%       1,
%       x
%     }
%   }
% };
% \path (3cm,-2cm) graph[tree layout] {
%   lam[as={$\lambda_{\texttt{x}}$}] -> {
%     * -> {
%       x,
%       2
%     }
%   }
% };
% \path (0cm,-4cm) graph[tree layout] {
%   lam[as={$\lambda_{\texttt{x}}$}] -> {
%     + -> {
%       1,
%       x
%     },
%     * -> {
%       x,
%       2
%     }
%   }
% };
% \path (2cm,-8cm) graph[tree layout] {
%   lam[as={$\lambda_{\texttt{x}}$}] -> {
%     + -> {
%       1,
%       * -> {
%         x,
%         2
%       }
%     }
%   }
% };
% \path (-2cm,-8cm) graph[tree layout] {
%   lam[as={$\lambda_{\texttt{x}}$}] -> {
%     * -> {
%       + -> {
%         1,
%         x
%       },
%       2
%     }
%   }
% };
\end{tikzpicture}
\caption{Conflicting wraps}
\label{fig:Conflicting wraps}
\end{figure}

If two different 
As seen in REF:SEC:TODO, conflicting wraps gener

Multi-parent

TODO: SHOW intermediate Steps (maybe no graph)
  Show user's perspective
  Accept one side = delete others (select this)
  Delete x in *
  Move + to x

\begin{figure}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[tree layout] {
  lam[as={$\lambda_{\texttt{x}}$}] -> {
    + -> {
      1,
      x
    },
    * -> {
      x,
      2
    }
  }
};
\path (2cm,-4cm) graph[tree layout] {
  lam[as={$\lambda_{\texttt{x}}$}] -> {
    + -> {
      1,
      * -> {
        x,
        2
      }
    }
  }
};
\path (-2cm,-4cm) graph[tree layout] {
  lam[as={$\lambda_{\texttt{x}}$}] -> {
    * -> {
      + -> {
        1,
        x
      },
      2
    }
  }
};
\end{tikzpicture}
\caption{Wrap Resolution}
\label{fig:Wrap Resolution}
\end{figure}

Say: "might not be a popup"

>>>>

% \subsubsection{Wrapping Code}

% \begin{figure}
% \centering
% TODO: naive add that creates a conflict
% \begin{tikzpicture}
% \path (0,0)       node {\includegraphics[scale=0.5]{graphviz/wrapping/1.dot}};
% \path (0cm,-4cm)  node {\includegraphics[scale=0.5]{graphviz/wrapping/2.dot}};
% \path (0cm,-8cm)  node {\includegraphics[scale=0.5]{graphviz/wrapping/3.dot}};
% \path (0cm,-12cm) node {\includegraphics[scale=0.5]{graphviz/wrapping/4.dot}};
% \path (0cm,-16cm) node {\includegraphics[scale=0.5]{graphviz/wrapping/5.dot}};
% \end{tikzpicture}
% \caption{Wrapping}
% \label{fig:Wrapping}
% \end{figure}

% Before \marginpar{DEL?} further exploring how to merge changes from multiple users,
% we want to point out a common idiom that we call \emph{wrapping}.
% Consider the first graph in Figure~\ref{fig:Wrapping}, which
% represents the code \texttt{$\lambda$ x . x}, and suppose
% the user's cursor is in the body field of the lambda (i.e., around the variable
% reference to \texttt{x} in the body of the lambda).

% If the user requests that a plus be inserted,
% the editor may notice that there is something already in that hole.
% A naive editor could certainly add the plus by creating a new
% edge from the body index of the lambda to a freshly created \texttt{plus} vertex.
% This would result in a conflict between the \texttt{plus} and the existing \texttt{var}
% as shown in Figure~REF:TODO.
% A better user experience
% would be interpret the request for a plus to mean to
% take whatever (if anything) is in the cursor
% and place it as a child of the plus.
% This \emph{wraps} the existing \texttt{var} in the \texttt{plus}
% and results in the code \texttt{$\lambda$ x . x + \hole}.
% (This is similar to the behavior in other structure editors such as in CITE:TODO.)
% Translating this into edge deletes and adds results in the sequence of graphs in Figure~\ref{fig:Wrapping}.
% Note that this is a layer above

% Traditional: no relation (and indent might change)

% \subsubsection{Conflicting wrap operations}

% \begin{figure}
% \centering
% \begin{tikzpicture}
% % TODO: texttt
% % TODO: position the different graphs
% \path (0,0)       node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/1.dot}};
% \path (-3cm,-3cm) node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/2A.dot}};
% \path (3cm,-3cm)  node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/2B.dot}};
% \path (0cm,-6cm)  node {\includegraphics[scale=0.5]{graphviz/conflicting-wraps/3.dot}};
% % \path (0cm,0cm) graph[tree layout] {
% %   lam[as={$\lambda_{\texttt{x}}$}] -> {
% %     x
% %   }
% % };
% % \path (-3cm,-2cm) graph[tree layout] {
% %   lam[as={$\lambda_{\texttt{x}}$}] -> {
% %     + -> {
% %       1,
% %       x
% %     }
% %   }
% % };
% % \path (3cm,-2cm) graph[tree layout] {
% %   lam[as={$\lambda_{\texttt{x}}$}] -> {
% %     * -> {
% %       x,
% %       2
% %     }
% %   }
% % };
% % \path (0cm,-4cm) graph[tree layout] {
% %   lam[as={$\lambda_{\texttt{x}}$}] -> {
% %     + -> {
% %       1,
% %       x
% %     },
% %     * -> {
% %       x,
% %       2
% %     }
% %   }
% % };
% % \path (2cm,-8cm) graph[tree layout] {
% %   lam[as={$\lambda_{\texttt{x}}$}] -> {
% %     + -> {
% %       1,
% %       * -> {
% %         x,
% %         2
% %       }
% %     }
% %   }
% % };
% % \path (-2cm,-8cm) graph[tree layout] {
% %   lam[as={$\lambda_{\texttt{x}}$}] -> {
% %     * -> {
% %       + -> {
% %         1,
% %         x
% %       },
% %       2
% %     }
% %   }
% % };
% \end{tikzpicture}
% \caption{Conflicting wraps}
% \label{fig:Conflicting wraps}
% \end{figure}

% If two different 
% As seen in REF:SEC:TODO, conflicting wraps gener

% Multi-parent

% TODO: SHOW intermediate Steps (maybe no graph)
%   Show user's perspective
%   Accept one side = delete others (select this)
%   Delete x in *
%   Move + to x

% \begin{figure}
% \centering
% \begin{tikzpicture}
% \path (0cm,0cm) graph[tree layout] {
%   lam[as={$\lambda_{\texttt{x}}$}] -> {
%     + -> {
%       1,
%       x
%     },
%     * -> {
%       x,
%       2
%     }
%   }
% };
% \path (2cm,-4cm) graph[tree layout] {
%   lam[as={$\lambda_{\texttt{x}}$}] -> {
%     + -> {
%       1,
%       * -> {
%         x,
%         2
%       }
%     }
%   }
% };
% \path (-2cm,-4cm) graph[tree layout] {
%   lam[as={$\lambda_{\texttt{x}}$}] -> {
%     * -> {
%       + -> {
%         1,
%         x
%       },
%       2
%     }
%   }
% };
% \end{tikzpicture}
% \caption{Wrap Resolution}
% \label{fig:Wrap Resolution}
% \end{figure}

% NEEDS UI MOCKUP

% Say: "might not be a popup"
